Dominance Regions for Rank Two Cluster Algebras

### 2022-04-04 dominance_regions.tex
Random intro edits

  Over time several bases have been constructed in a variety of generalities \cite{foo,bar,baz}. 
  All of them share a property of being ``pointed''.
  Being pointed turned out to be a desireable feature for a basis to have and a natural question is to find all bases enjoying this property.
  Recently Qin introduced the notion of \emph{dominance} and showed that it completely encapsulates the deformability of pointed bases whenever the cluster algebra has full rank \cite{Qin}. 
  As no explicit calculation is carried out in his work, we aim here at describing explicitly the space of deformability of pointed bases in rank 2, i.e. when clusters contain a pair of mutable cluster variables.
  In this setting, frozen variables do not carry any additional information and we can work in the coefficient-free case.

  Fix integers $b,c>0$.  The cluster algebra $\cA(b,c)$ is the $\ZZ$-subalgebra of $\QQ(x_1,x_2)$ generated by \emph{cluster variables} $x_m$, $m\in\ZZ$, defined recursively by
  \begin{equation}
    x_{m-1}x_{m+1}=\begin{cases} x_m^b+1 & \text{if $m$ is odd;}\\ x_m^c+1 & \text{if $m$ is even.} \end{cases}
  \end{equation}
  By the Laurent Phenomenon \cite{fomin-zelevinsky}, each $x_m$ is actually an element of $\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for any $k\in\ZZ$.
  Moreover, these Laurent polynomials are known to have positive coefficients \cite{lee-schiffler,GHKK}.

  Our goal in this note is to explicitly compute the possible bases which are \emph{pointed} with respect to the embedding $\cA(b,c)\into\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for every $k\in\ZZ$.
  That is, a basis element, with \emph{$\bfg$-vector} $\lambda\in\ZZ^2$ when expanded in terms of $\{x_1,x_2\}$, can be written in the form
  \[
    x_1^{\lambda_1}x_2^{\lambda_2}\sum\limits_{a_1,a_2 \ge 0} c^\lambda_{a_1,a_2} x_1^{-ba_1} x_2^{ca_2}
  \]
  with $c_{0,0}=1$ and an analogous structure reproduces after expanding in terms of any pair $\{x_k,x_{k+1}\}$.
  %The collection of all $\bfg$-vectors $\lambda\in\ZZ^2$ parametrize the elements of any pointed basis.

@@

  Our calculations reveal the maximal support possible for a pointed basis which remains pointed under all possible mutations.
  In particular, any candidate basis can be ruled out using these results based upon their support alone.

@@

>>>>

  Over time several bases for cluster algebras have been constructed in a variety of generalities \cite{foo,bar,baz}. 
  All of them share a property of being ``pointed''.
  Being pointed turned out to be a desirable feature for a basis to have and a natural question is to find all bases enjoying this property.
  Recently Qin introduced the notion of \emph{dominance} and showed that it completely encapsulates the deformability of pointed bases whenever the cluster algebra has full rank \cite{Qin}. 
  As no explicit calculation is carried out in his work, we aim here at describing explicitly the space of deformability for pointed bases in rank 2, i.e. when clusters contain a pair of mutable cluster variables.
  In this setting, frozen variables do not carry any additional information and we can work in the coefficient-free case.

  Fix integers $b,c>0$.  The cluster algebra $\cA(b,c)$ is the $\ZZ$-subalgebra of $\QQ(x_1,x_2)$ generated by \emph{cluster variables} $x_m$, $m\in\ZZ$, defined recursively by
  \begin{equation}
    x_{m-1}x_{m+1}=\begin{cases} x_m^b+1 & \text{if $m$ is odd;}\\ x_m^c+1 & \text{if $m$ is even.} \end{cases}
  \end{equation}
  By the Laurent Phenomenon \cite{fomin-zelevinsky}, each $x_m$ is actually an element of $\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for any $k\in\ZZ$.
  Moreover, these Laurent polynomials are known to have positive coefficients \cite{LLZ,lee-schiffler,GHKK}.

  Our goal in this note is to explicitly compute the possible bases which are \emph{pointed} with respect to the embedding $\cA(b,c)\into\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for every $k\in\ZZ$.
  That is, a basis element with \emph{$\bfg$-vector} $\lambda\in\ZZ^2$ when expanded in terms of $\{x_1,x_2\}$ can be written in the form
  \[
    x_1^{\lambda_1}x_2^{\lambda_2}\sum\limits_{a_1,a_2 \ge 0} c^\lambda_{a_1,a_2} x_1^{-ba_1} x_2^{ca_2}
  \]
  with $c^\lambda_{0,0}=1$ and an analogous structure reproduces after expanding in terms of any pair $\{x_k,x_{k+1}\}$.
  %The collection of all $\bfg$-vectors $\lambda\in\ZZ^2$ parametrize the elements of any pointed basis.

@@

@@

\begin{remark}
  This result reveals the maximal support possible for a pointed basis which remains pointed under all possible mutations.
  In particular, any candidate basis could be ruled out using these results based upon their support alone.
\end{remark}
